---
title: "chp2"
author: "Vasco Brazão"
date: "11/14/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages
```{r packages}
library(tidyverse)
```

## 2E1

*The probability of rain on Monday* could be restated as *the probability of rain, given that it is Monday*: Pr(Rain|Monday), answer (2)

## 2E2

Answer (3).

## 2E3 

Answer (1).

## 2E4

The statement can be interpreted in many ways. For instance, we could interpret it to mean that the 70% globe surface is colored blue, and that any of the points on the surface is equally likely to land under your right index finger.

## 2M1

Computing and plotting the grid approximate posterior distribution for different sets of observations. I will use 20 bins for this, and do my best to use tidyverse. Let's see how it goes.

(1) W, W, W
(2) W, W, W, L
(3) L, W, W, L, W, W, W

```{r 2m1.grid}
p_grid <- tibble(
  # define the grid
  probs = seq(from = 0, to = 1, length.out = 20),
  # define the prior
  prior = 1
)
```

Above I defined the grid. Now I would need to compute the likelihood, for each probability, of the data.Then multiply the likelihood with the prior, and then standardize. 

```{r 2m1.posteriors}
p_grid <- p_grid %>%
  mutate(
    # likelihood for (1), (2), then (3)
    first = dbinom(3, size = 3, prob = probs),
    secon = dbinom(3, size = 4, prob = probs),
    third = dbinom(5, size = 7, prob = probs),
    
    # posteriors
    first_uns_pos = first * prior,
    secon_uns_pos = secon * prior,
    third_uns_pos = third * prior,
    
    # standardizing
    first_pos = first_uns_pos / sum(first_uns_pos),
    secon_pos = secon_uns_pos / sum(secon_uns_pos),
    third_pos = third_uns_pos / sum(third_uns_pos)
  )
```

Now plots!

```{r 2m1.plots}
# graph for (1)
p_grid %>%
  ggplot(aes(x = probs, y = first_pos)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Grid approx. posterior distribution for probability of W",
    subtitle = "Data = W, W, W",
    x = "Probability of W",
    y = "Posterior"
  ) +
  theme_bw()

# graph for (2)
p_grid %>%
  ggplot(aes(x = probs, y = secon_pos)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Grid approx. posterior distribution for probability of W",
    subtitle = "Data = W, W, W, L",
    x = "Probability of W",
    y = "Posterior"
  ) +
  theme_bw()

# graph for (3)
p_grid %>%
  ggplot(aes(x = probs, y = third_pos)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Grid approx. posterior distribution for probability of W",
    subtitle = "Data = L, W, W, L, W, W, W",
    x = "Probability of W",
    y = "Posterior"
  ) +
  theme_bw()
```

Yuhu!

Some thoughts: lots of repetitions in my code. If I try to improve it later, I might:

* Create actual variables for the data, and then a function that extracts the correct values to place in dbinom() 
* Create a function (or loop??) to generate the plots without repeating myself so much. 
