---
title: "chp6"
author: "Vasco Brazão"
date: "09/06/2021"
output: pdf_document
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
library(tidyverse)
library(here)
```

I did 6E1 - 6M1 in a paper notebook and am too lazy to type things up right now.

## 6M2

We should simulate from DAG X -> Z -> Y, such that the correlation between X and Z is very large.

```{r 6m2.data}
Nsim <- 1000

dat <- tibble::tibble(
  x = rnorm(n = Nsim),
  z = rnorm(n = Nsim, mean = x, sd = 0.5),
  e = rnorm(n = Nsim),
  y = rnorm(n = Nsim, mean = z + e)
)

pairs(dat)

cor(dat)
```
```{r 6m2.lm}
summary(lm(y ~ x + z, data = dat))
```
First, we can see that we were successful in simulating from the DAG and creating a rather large association between the predictors.

Next, we see no "problem" in the linear model results -- only z has a a large estimated coefficient. Let's replicate this result in `quap` and `brms`.

```{r 6m2.quap}
q6m2 <- rethinking::quap(
  alist(
    y ~ dnorm(mu, sigma),
    mu <- a + bx*x + bz*z,
    a ~ dnorm(0, 2),
    bx ~ dnorm(0, 2),
    bz ~ dnorm(0, 2),
    sigma ~ dexp(1)
  ), data = dat
)

rethinking::precis(q6m2)

library(rethinking)

plot(precis(q6m2))

post <- rethinking::extract.samples(q6m2)

plot(bx ~ bz, post, col = col.alpha(rangi2, 0.1), pch = 16)

cor(post$bx, post$bz)
```
Here we see that, conditional on the model, bx and bz are quite strongly correlated.

The diference here is that X does not actually affect Y directly, while in the legs example both legs are affected by the height of the person.

Let's just code this model in brms.

```{r 6m3.brms}
detach(package:rethinking, unload=TRUE)

library(brms)

b6m2 <- brm(
  data = dat,
  family = gaussian,
  y ~ 1 + x + z,
  prior = c(prior(normal(0, 2), class = Intercept),
            prior(normal(0, 2), class = b),
            prior(exponential(1), class = sigma)),
  iter = 2000, warmup = 1000, chains = 4, cores = 4,
  file = here::here("chp6", "b6m2")
)

print(b6m2)

brms::mcmc_plot(b6m2,
               type = "intervals",
               prob = .5,
               prob_outer = .95,
               point_est = "mean")

pairs(b6m2, pars = parnames(b6m2)[2:3])
```
Ooh what cool plots we can make.

## 6M3

Going clockwise, starting at top left DAG:

1. Z 
2. nothing
3. nothing
4. A

Needed a lot of help from wjakethompson. Perhaps I'm a bit rusty. Totally forgot that we want to include indirect causal paths.


