---
title: "chp4"
author: "Vasco Brazão"
date: "12/23/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages
```{r packages}
library(tidyverse)
library(brms)
library(here)
```

## 4E1

The first line is the likelihood, the two others lines are the priors for mu and sigma.

## 4E2

2 parameters, mu and sigma

## 4E3

Oh god. Still need to learn how to format things so...

Pr(mu, sigma|y_i) = PRODUCT_i(Normal(y_i|mu, sigma) * Normal(mu|0, 10) * Exponential(sigma|1) / DOUBLE_INTEGRAL(PRODUCT_i(Normal(y_i|mu, sigma) * Normal(mu|0, 10) * Exponential(sigma|1) )) dmu dsigma

Happy?

## 4E4

The second line is the linear model.

## 4E5

There are three parameters: alpha, beta, and sigma.

## 4M1

```{r 4m1.sim}
N <- 100000

y_values <- tibble(
  mu = rnorm(N, 0, 10),
  sigma = rexp(N, 1),
  y = rnorm(N, mean = mu, sd = sigma)
)


hist(y_values$y, breaks = 1000)
```

## 4M2

As I'm using `brms`, and trying to learn how to use it at the same time, I'll skip the `quap` stuff and jump straight into `brms`.

```{r 4m2}
#this is what it maybe would look like in brms. hell if I know

b4m1 <- brm(data = y_values,
            family = gaussian,
            y ~ 1,
            prior = c(prior(normal(0, 10), class = Intercept),
                      prior(exponential(1), class = sigma)),
            iter = 2000, warmup = 1000, chains = 4, cores = 4,
            seed = 4,
            file = here("b4m1")
)
```

```{r 4m2.chains}

plot(b4m1)

```

## 4M4

$$
\begin{align*}
\text{height}_i & \sim \operatorname{Normal}(\mu_i, \sigma) \\
\mu_i  & = \alpha + \beta \times (\text{year}_i - 2) \\
\alpha & \sim \operatorname{Normal}(160, 10) \\
\beta  & \sim \operatorname{Normal}(0, 5) \\
\sigma & \sim \operatorname{Exponential}(1/10) \\
\end{align*}
$$

## 4M5

Of course, I almost forgot!

$$
\begin{align*}
\text{height}_i & \sim \operatorname{Normal}(\mu_i, \sigma) \\
\mu_i  & = \alpha + \beta \times (\text{year}_i - 2) \\
\alpha & \sim \operatorname{Normal}(140, 20) \\
\beta  & \sim \operatorname{Log-Normal}(0, 5) \\
\sigma & \sim \operatorname{Exponential}(1/10) \\
\end{align*}
$$

## 4M6 

Assuming also that our sample has kids of roughly the same age, we can adjust the prior for the variance. But not sure this one makes sense?

$$
\begin{align*}
\text{height}_i & \sim \operatorname{Normal}(\mu_i, \sigma) \\
\mu_i  & = \alpha + \beta \times (\text{year}_i - 2) \\
\alpha & \sim \operatorname{Normal}(140, 32) \\
\beta  & \sim \operatorname{Log-Normal}(0, 5) \\
\sigma & \sim \operatorname{Exponential}(1/32) \\
\end{align*}
$$