---
title: "chp3"
author: "Vasco Brazão"
date: "12/5/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages
```{r packages}
library(tidyverse)
library(rethinking)
```

Code for the easy problems (from book):

```{r easy.samples}

p_grid <- seq( from=0 , to=1 , length.out=1000 )
prior <- rep( 1 , 1000 )
likelihood <- dbinom( 6 , size=9 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)

set.seed(100)

samples <- sample(p_grid, prob = posterior, size = 1e4, replace = TRUE)
```

## 3E1

```{r 3e1}
sum(samples < .2) / length(samples)
```
## 3E2

```{r 3e2}
sum(samples > .8) / length(samples)
```
## 3E3

```{r 3e3}
(sum(samples > .2) - sum(samples >= .8))/ length(samples)
```
## 3E4

```{r 3e4}
quantile(samples, .2)
```
## 3E5

```{r 3e5}
quantile(samples, .8)
```
## 3E6

```{r 3e6}
rethinking::HPDI(samples, .66)
```

## 3E7

```{r 3e7}
rethinking::PI(samples, .66)
```
## 3M1
This time trying to be a little more tidyverse-y
```{r 3m1}
globe <- tibble(
  p_grid = seq(from = 0, to = 1, length.out = 1000),
  prior = rep(1, 1000),
  likelihood = dbinom(8, size = 15, prob = p_grid),
  posterior = likelihood * prior / sum(likelihood * prior)
)

#a graph for good measure

globe3m1post <- globe %>%
  ggplot(aes(x = p_grid, y = posterior)) +
  geom_line()

print(globe3m1post)
```

## 3M2

```{r 3m2}
set.seed(100)

samples <- sample(globe$p_grid, prob = globe$posterior, size = 1e4, replace = TRUE)

rethinking::HPDI(samples, .90)
```

## 3M3

```{r 3m3}
w <- rbinom(1e4, size = 15, prob=samples)

# probability of 8 W
sum(w == 8)/1e4
```
## 3M4

```{r 3m4}
new_w <- rbinom(1e4, size = 9, prob = samples)

# probability of 6 w
sum(new_w == 6)/1e4

```

## 3M5
```{r 3m5}
# first, we create the data again
globe_new <- tibble(
  p_grid = seq(from = 0, to = 1, length.out = 1000),
  prior = ifelse(p_grid >= .5, 1, 0),
  likelihood = dbinom(8, size = 15, prob = p_grid),
  posterior = likelihood * prior / sum(likelihood * prior)
)

# graph

globe3m5post <- globe_new %>%
  ggplot(aes(x = p_grid, y = posterior)) +
  geom_line()

print(globe3m5post)
```
To calculate de hpdi

```{r 3m5.hpdi}
set.seed(100)

samples <- sample(globe_new$p_grid, prob = globe_new$posterior, size = 1e4, replace = TRUE)

rethinking::HPDI(samples, .90)
```

We can see it is quite a bit narrower, and now confidently excludes values below .5. The upper bound did not shrink by a lot though.

For the posterior predictive check: what does our model think is the probability of observing 8 waters in 15 tosses?

```{r 3m5.ppc}
w <- rbinom(1e4, size = 15, prob = samples)

# probability of 8 W
sum(w == 8)/1e4
```

It is a little lower than before.

Now for 6 waters in 9 tosses.

```{r 3m5.sixwaters}
new_w <- rbinom(1e4, size = 9, prob = samples)

# probability of 6 w
sum(new_w == 6)/1e4
```

It is a little higher than before.

All in all, the model performs a bit better.

## 3M6

This one looks fun! Not sure if there is an easy way to do it, but I will try some brute force I suppose. Will also help me brush up on functions and simulations and whatnot, I guess.

```{r 3m6}
pi_width <- function(tosses = 20){
  # the amount of waters we observe is random, but depends on the true 
  # proportion which we take to be .7
  waters <- rbinom(1, size = tosses, prob = .7)
  
  # now we can construct the posterior, using first the flat prior
  globe_pos <- tibble(
    p_grid = seq(from = 0, to = 1, length.out = 1000),
    prior = rep(1, 1000),
    likelihood = dbinom(waters, size = tosses, prob = p_grid),
    posterior = likelihood * prior / sum(likelihood * prior)
  )
  
  # then we sample from the posterior
  pos_samples <- sample(globe_pos$p_grid, 
                        prob = globe_pos$posterior, 
                        size = 1e4, replace = TRUE)
  
  # now we feed the samples to the PI function
  interval <- rethinking::PI(pos_samples, .99)
  
  # this is what we want to minimize, what our function will return
  width <- interval[[2]] - interval[[1]]
  
  return(width)
}

# I want to get 10 data points for different values of tosses, and just see what
# that looks like. After some experiment, we narrow in on the range 2000:2500

width_data <- 
  tibble(
    n_tosses = rep(2000:2500, each = 10),
    width = map_dbl(.x = n_tosses, .f = ~ pi_width(tosses = .x))
  )


width_data %>%
  ggplot(aes(x = n_tosses, y = width)) +
  geom_point(alpha = .5) +
  geom_smooth()

width_data %>%
  group_by(n_tosses) %>%
  mutate(
    avg = mean(width)
  ) %>%
  filter(avg <= .05) %>%
  select(n_tosses, avg) %>%
  distinct()
```

If we throw the globe more than 2170, the expected width of the interval is less than .05. 

Wonder how other people solved this!

## 3H1

```{r hard.data}
births <- tibble(
  birth1 = c(1,0,0,0,1,1,0,1,0,1,0,0,1,1,0,1,1,0,0,0,1,0,0,0,1,0,
0,0,0,1,1,1,0,1,0,1,1,1,0,1,0,1,1,0,1,0,0,1,1,0,1,0,0,0,0,0,0,0,
1,1,0,1,0,0,1,0,0,0,1,0,0,1,1,1,1,0,1,0,1,1,1,1,1,0,0,1,0,1,1,0,
1,0,1,1,1,0,1,1,1,1),
  birth2 = c(0,1,0,1,0,1,1,1,0,0,1,1,1,1,1,0,0,1,1,1,0,0,1,1,1,0,
1,1,1,0,1,1,1,0,1,0,0,1,1,1,1,0,0,1,0,1,1,1,1,1,1,1,1,1,1,1,1,1,
1,1,1,0,1,1,0,1,1,0,1,1,1,0,0,0,0,0,0,1,0,0,0,1,1,0,0,1,0,0,1,1,
0,0,0,1,1,1,0,0,0,0)
  )

n_boys <- sum(births$birth1) + sum(births$birth2)
n_births <- length(births$birth1) + length(births$birth2)

grid_births <- tibble(
  prob = seq(from = 0, to = 1, length.out = 1000),
  prior = rep(1, 1000),
  likelihood = dbinom(n_boys, size = n_births, prob = prob),
  posterior = likelihood * prior / sum(likelihood * prior)
)



grid_births[which.max(grid_births$posterior), ]$prob
```
## 3H2

```{r 3h2}
births_sample <- sample(grid_births$prob, 
                        size = 1e4, 
                        prob = grid_births$posterior,
                        replace = TRUE)

for (i in c(.5, .89, .97)){
  print(rethinking::HPDI(births_sample, i))
}

```

## 3H3

```{r 3h3}
replicates <- tibble(
  boys = rbinom(n = 1e4,
                size = 200, 
                prob = births_sample)
)

replicates %>%
  ggplot(aes(x = boys)) +
  geom_density() +
  geom_vline(xintercept = 111)

```

The distribution seems ok. The most likely outcome is the "correct" outcome. 

## 3H4

```{r 3h4}
replicates_first <- tibble(
  boys_first = rbinom(n = 1e4,
                size = 100, 
                prob = births_sample)
)

replicates_first %>%
  ggplot(aes(x = boys_first)) +
  geom_density() +
  geom_vline(xintercept = sum(births$birth1))
```

Now the most likely outcome seems quite far away from the true value.

## 3H5

```{r 3h5}
girls_first <- length(births$birth1) - sum(births$birth1)

boys_after_girls <- births %>%
  filter(birth1 == 0) %>%
  summarise(
    boys = sum(birth2)
  )

replicates_first_girls <- tibble(
  boys_after_girls = rbinom(n = 1e4,
                size = 49, 
                prob = births_sample)
)

replicates_first_girls %>%
  ggplot(aes(x = boys_after_girls)) +
  geom_density() +
  geom_vline(xintercept = boys_after_girls[[1]])
```

It seems there were many more boys born after a girl than our model (based on the total data) would predict. But I'm not sure why?